Assessing the Various Influences on Current Coffee Prices

### 2024-12-15 paper/paper.qmd
Update context of paper
Delete useless word or snetences and add useful information.

In this study of coffee pricing strategies in Canadian grocery stores, the focus is on estimating how historical prices (old price), time of year (month), and vendor characteristics (vendor) affect current coffee prices. Using a Bayesian multiple linear regression model, I estimated these effects to better understand the pricing dynamics of coffee products and to determine how much each factor contributes to pricing decisions in the market. This modeling approach provides a probabilistic framework that accounts for uncertainty in the estimates, allowing for more nuanced conclusions compared to traditional methods.

The findings reveal that the old price has the most significant positive impact on current coffee prices among all the predictors considered. This suggests that vendors heavily rely on historical prices when determining current prices, possibly to maintain a level of consistency that consumers expect. Consistency in pricing is often valued by consumers, as it fosters a sense of trust and predictability. Due to a lack of data for some months, it was challenging to establish a definitive analysis of seasonal trends, though this aspect remains an important factor for understanding pricing behavior comprehensively. Seasonal patterns can have substantial effects on consumer behavior, and thus their absence in the dataset may limit the scope of understanding regarding periodic fluctuations in coffee prices.

@@

Since the variable nowtime only records 4 months, it is considered a lack of Long-Term Trends, which means it's difficult to identify long-term pricing or demand patterns by using short data periods. So I only extracted a new variable—month from date of nowtime, which can simplify temporal analysis and identify trends, such as seasonal monthly pattern with price changes or demand patterns. It allows grouping data for monthly aggregation and supporting seasonality-focused understanding or forecasting models.

@@

The dataset from Hammer provides a wild representation of real-world retail activities, capturing key elements such as product details, vendor listings, and price updates. When vendors make updates to product information, including pricing changes or availability adjustments, Hammer gathers and organizes this data into a structured dataset. This ensures that the dataset remains accurate and reflective of current retail conditions.

The dataset includes essential fields such as vendor, product name, current price, old price, and a timestamp referred to as nowtime. The data is collected through scraping and is systematically structured to support an in-depth analysis of retail trends, pricing strategies, and overall market dynamics across different periods. Each record captures a snapshot of a product’s market presence at a given time, enabling targeted analyses like monitoring price fluctuations for specific products, such as coffee. By offering such detailed and time-specific information, the dataset provides a important foundation for understanding the evolving dynamics of retail pricing and vendor strategies.

@@

The predictor variables selected from the raw dataset are old price, month, and vendor. The old price serves as a predictor because it provides beneficial understanding into how past pricing decisions influence current prices. Vendors often make incremental adjustments based on historical pricing data, consumer behavior, and market reactions, which makes it a key factor in understanding pricing strategies. Month is included to account for possible seasonal trends in pricing, as consumer demand and promotional activities often vary throughout the year, potentially impacting coffee prices. Lastly, vendor is used to examine pricing variations between Metro and SaveOnFoods. Each vendor may have distinct pricing approaches based on factors like target market, cost structures, or regional positioning, and including this variable enables a comparative analysis of vendor-specific pricing dynamics.

>>>>

In this study of coffee pricing strategies in Canadian grocery stores, the focus is on estimating how historical prices (old price), time of year (month), and vendor characteristics (vendor) affect current coffee prices. Using a Bayesian multiple linear regression model, I estimated these effects to better understand the pricing dynamics of coffee products and to determine how much each factor contributes to pricing decisions in the market. This modeling approach provides a probabilistic framework that accounts for uncertainty in the estimates. And it allows for more detailed conclusions compared to normal methods.

The findings reveal that the old price has the most significant positive impact on current coffee prices among all the predictors considered. This suggests that vendors heavily rely on historical prices when determining current prices, possibly to maintain a level of consistency that consumers expect. Consistency in pricing is often valued by consumers, as it fosters a sense of trust and predictability. Due to a lack of data for some months, it was challenging to establish a definitive analysis of seasonal trends, though this aspect remains an important factor for understanding pricing behavior comprehensively. Time variable can have substantial effects on consumer behavior, and thus their absence in the dataset may limit the scope of understanding regarding periodic fluctuations in coffee prices.

@@

Since the variable nowtime only records 4 months, it is considered a lack of Long-Term Trends, which means it's difficult to identify long-term pricing or demand patterns by using short data periods. So I only extracted a new variable—month from date of nowtime, which can figure out the problem of lacking data of time variable.

@@

The dataset from Hammer provides a representation of real-world retail activities, includingg key elements such as product details, vendor listings, and price updates currently and historically. When vendors make updates to product information, including pricing changes or availability adjustments, Hammer gathers and organizes this data into a structured dataset. This ensures that the dataset remains accurance and reflective of current retail conditions.

The dataset includes vendor, product name, current price, old price, and a times variable referred to nowtime. The data is collected through scraping and is systematically structured to support an in-depth analysis of retail trends, potential pricing relationships, and overall market dynamics across different periods. Each record captures a snapshot of a product’s market presence at a given time, enabling targeted analyses like monitoring fluctuation between current and historical price for specific products, such as coffee which is the product type the paper wants to analyze. By offering such detailed and time-specific information, the dataset provides a important foundation for understanding the evolving dynamics of retail pricing and potential impact.

@@

The predictor variables selected from the raw dataset are old price, month, and vendor. The old price serves as a predictor because it provides how past pricing decisions influence current prices. Vendors often make incremental adjustments based on historical pricing data, consumer behavior, and market reactions, which makes it a key factor in understanding why price changing in some way. Month is included to account for possible seasonal trends in pricing, as consumer demand and promotional activities often vary throughout the year, potentially impacting coffee prices. Lastly, vendor is used to examine pricing variations between Metro and SaveOnFoods. Each vendor may have distinct pricing approaches based on factors like target market, cost structures, or regional positioning, and including this variable enables a comparative analysis of vendor-specific pricing dynamics.
